---
title: "Homework 10"
author: "Shelby Nechkash, Seana, Greg"
date: "2023-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load accident data and plot latitude and longitude on a map

```{r}
library(tidyverse)
accidentdata = Crash_Data
ggplot(accidentdata, aes(x=LONGITUD, y=LATITUDE)) + geom_point()
#need to adjust where the axis starts
```


2. Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

```{r}
accidentdata %>% group_by(DAY_WEEK) 
#need to finish to count
```


3.What time of the day do accidents happen (use variable HOUR)?


```{r}
accidentdata %>% group_by(HOUR)
#Need to add a count
```

4.What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
accidentdata %>% filter(DRUNK_DR >=1)
#need to add count
```


5.Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.

6.Join accident and driver table (work out which variable(s) to use)

7.Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!

