---
title: "Homework 10"
author: "Shelby Nechkash, Seana, Greg"
date: "2023-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load accident data and plot latitude and longitude on a map

```{r}
library(tidyverse)
library(maps)
accidentdata = read_csv("accident data.csv");
#ggplot(accidentdata, aes(x=LONGITUD, y=LATITUDE)) + geom_point()

states <- map_data("state")
head(states)

#states %>% ggplot(aes(x=long, y=lat)) + geom_path(aes(group=group)) + geom_point(aes(x=LONGITUD, y=LATITUDE), size=0.05, color='red', data=acc %>% filter(LONGITUDE<0, LLONGUDE>-180))

#ggplot(map(aes(x=long, y=lat))) + geom_polygon(aes(group=group))  + Geom_point(accidentdata, aes(x=LONGITUD, y=LATITUDE))
```


2. Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

Most accidents occur on Friday, Saturday, and Sunday

```{r}
accidentdata %>% group_by(DAY_WEEK) %>% mutate(DAY_WEEK = wday(DAY_WEEK, label=TRUE)) %>% summarise(n = n())
```


3.What time of the day do accidents happen (use variable HOUR)?


Most accidents occur between 17:00 and 20:00.

```{r, warning=FALSE}
accidentdata %>% group_by(HOUR) %>% mutate(DAY_WEEK = hours(HOUR)) %>% summarise(n = n()) 

```

4.What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

There is a total of 8720 accidents with at least one drunk driver. 

```{r}
accidentdata %>% filter(DRUNK_DR >=1) %>% summarise(n = n()) 

```


5.Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.

6.Join accident and driver table (work out which variable(s) to use)

7.Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!

