---
title: "Homework 10"
author: "Shelby Nechkash, Seana, Greg"
date: "2023-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Repository Link - https://github.com/shelbob/Homework10 

1. Load accident data 

```{r}
library(tidyverse)
accidentdata = read_csv("accident data.csv");

```


2. Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

Most accidents occur on Friday, Saturday, and Sunday

```{r}
accidentdata %>% group_by(DAY_WEEK) %>% mutate(DAY_WEEK = wday(DAY_WEEK, label=TRUE)) %>% summarise(n = n())
```


3.What time of the day do accidents happen (use variable HOUR)?


Most accidents occur between 17:00 and 20:00.

```{r, warning=FALSE}
accidentdata %>% group_by(HOUR) %>% mutate(DAY_WEEK = hours(HOUR)) %>% summarise(n = n()) 

```

4.What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

There is a total of 8720 accidents with at least one drunk driver. 

```{r}
accidentdata %>% filter(DRUNK_DR >=1) %>% summarise(n = n()) 

```


5.Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.

```{r}

person <- read.csv("https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/person.csv")
str(person)
person$PER_TYP <- as.factor(person$PER_TYP)

drivers <- person %>% filter(PER_TYP == 1)
summary(drivers$PER_TYP)
```

6.Join accident and driver table (work out which variable(s) to use)

```{r}

join.data <- drivers %>% left_join(accidentdata, by="ST_CASE")

```


7.Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!

```{r}

join.data$DAY_WEEK <- as.factor(join.data$DAY_WEEK)
join.data$HOUR.x <- as.factor(join.data$HOUR.x)
join.data$SEX <- as.factor(join.data$SEX)
summary(join.data$DAY_WEEK)
summary(join.data$HOUR.x)
summary(join.data$SEX)

ggplot(join.data, aes(x = DAY_WEEK)) + geom_bar()
ggplot(join.data, aes(x = HOUR.x)) + geom_bar()
ggplot(join.data, aes(x = SEX)) + geom_bar()
```
